---
title: "DATA 621: BUSINESS ANALYTICS AND DATA MINING HOMEWORK#4: LOGISTIC REGRESSION"
date: "Last edited `r format(Sys.time(), '%B %d, %Y')`"
author: "Group 2 - Gabriel Campos, Melissa Bowman, Alexander Khaykin, & Jennifer Abinette"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 4
    number_sections: true
    highlight: tango
  geometry: "left=0.5cm,right=0.5cm,top=0.5cm,bottom=0.5cm"
  html_document:
    df_print: paged
urlcolor: blue
---

# Overview

&emsp; In this homework assignment, you will explore, analyze and model a data set containing approximately 8000
records representing a customer at an auto insurance company. Each record has two response variables. The
first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero
means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero
if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.

&emsp; Your objective is to build multiple linear regression and binary logistic regression models on the training data
to predict the probability that a person will crash their car and also the amount of money it will cost if the person
does crash their car. You can only use the variables given to you (or variables that you derive from the variables
provided). Below is a short description of the variables of interest in the data set:
<br>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library('dplyr')
library('kableExtra')
library(tidyverse)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(psych)
library(car)
library(caret)
options(scipen = 999)
```

```{r, echo=FALSE}
ls_c1 <-c("INDEX", "TARGET_FLAG", "TARGET_AMT", "AGE", "BLUEBOOK", "CAR_AGE",
             "CAR_TYPE", "CAR_USE", "CLM_FREQ #", "EDUCATION", "HOMEKIDS #",
             "HOME_VAL", "INCOME", "JOB","KIDSDRIV #", "MSTATUS", "MVR_PTS",
             "OLDCLAIM", "PARENT1", "RED_CAR", "REVOKED", "SEX Gender", "TIF",
             "TRAVTIME", "URBANICITY", "YOJ")

ls_c2 <- c("Identification Variable (do not use)",
              "Was Car in a crash? 1=YES 0=NO",
              "If car was in a crash, what was the cost",
              "Age of Driver", "Value of Vehicle", "Vehicle Age", "Type of Car",
              "Vehicle Use", "Claims (Past 5 Years)", "Max Education Level",
              "Children at Home", "Home Value", "Income",
              "Category", "Driving Children", "Marital Status",
              "Motor Vehicle Record Points", "Total Claims (Past 5 Years)",
              "Single Parent", "A Red Car", "License Revoked (Past 7 Years)",
              "Gender", "Time in Force", "Distance to Work", "Home/Work Area",
              "Years on Job")

ls_c3 <- c("None", "None", "None",
             "Very young people tend to be risky. Maybe very old people also.",
             paste("Unknown effect on probability of collision, but probably ",
              "effect the payout if there is a crash"),
             paste("Unknown effect on probability of collision, but probably ",
              "effect the payout if there is a crash"),
            paste("Unknown effect on probability of collision, but probably ",
              "effect the payout if there is a crash"),
             paste("Commercial vehicles are driven more, so might ",
              "increase probability of collision"),
             paste("The more claims you filed in the past, the more you are ",
              "likely to file in the future"),
             paste("Unknown effect, but in theory more educated people ",
              "tend to drive more safely"),
             "Unknown effect", 
            "In theory, home owners tend to drive more responsibly",
             "In theory, rich people tend to get into fewer crashes",
             "In theory, white collar jobs tend to be safer",
             paste("When teenagers drive your car, you are more likely to",
              " get into crashes"),
             "In theory, married people drive more safely",
             paste("If you get lots of traffic tickets, you tend to get",
              " into more crashes"),
             paste("If your total payout over the past five years ",
               "was high, this suggests future payouts will be high"),
             "Unknown effect",
             paste("Urban legend says that red cars (especially red ",
             "sports cars) are more risky. Is that true?"),
             paste("If your license was revoked in the past 7 years, ",
             "you probably are a more risky driver."),
             paste("Urban legend says that women have less crashes",
             " then men. Is that true?"),
             paste("People who have been customers for a long time",
             " are usually more safe."),
             "Long drives to work usually suggest greater risk",
             "Unknown",
             "People who stay at a job for a long time are usually more safe")  
  
df_overview <- data.frame(column1 = ls_c1,
                            Definition = ls_c2,
                            column3 = ls_c3)%>%
                              rename(`Variable Names` = column1, 
                                     `Theoretical Effect` = column3)

df_overview%>%
  kable()%>%
    kable_styling(latex_options = c('scale_down','HOLD_position'))%>%
      column_spec(3, width = "20em")
```

```{r, echo=FALSE}
rm(ls_c1,ls_c2,ls_c3,df_overview)
```

\

## Deliverables

* A write-up submitted in PDF format. Your write-up should have four sections. Each one is described below. You may assume you are addressing me as a fellow data scientist, so do not need to shy away from technical details.
* Assigned predictions (probabilities, classifications, cost) for the evaluation data set. Use 0.5 threshold.
* Include your R statistical programming code in an Appendix.

## Write Up:

### 1. DATA EXPLORATION (25 Points)

Describe the size and the variables in the insurance training data set. Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job. Some suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment. You should have your own thoughts on what to tell the boss. These are just ideas.

a. Mean / Standard Deviation / Median
b. Bar Chart or Box Plot of the data
c. Is the data correlated to the target variable (or to other variables?)
d. Are any of the variables missing and need to be imputed “fixed”?

### 2. DATA PREPARATION (25 Points)

Describe how you have transformed the data by changing the original variables or creating new variables. If you did transform the data or create new variables, discuss why you did this. Here are some possible transformations.

a. Fix missing values (maybe with a Mean or Median value)
b. Create flags to suggest if a variable was missing
c. Transform data by putting it into buckets
d. Mathematical transforms such as log or square root (or use Box-Cox)
e. Combine variables (such as ratios or adding or multiplying) to create new variables

### 3. BUILD MODELS (25 Points)

Using the training data set, build at least two different multiple linear regression models and three different binary logistic regression models, using different variables (or the same variables with different transformations). You may select the variables manually, use an approach such as Forward or Stepwise, use a different approach such as trees, or use a combination of techniques. Describe the techniques you used. If you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done. 


Discuss the coefficients in the models, do they make sense? For example, if a person has a lot of traffic tickets, you would reasonably expect that person to have more car crashes. If the coefficient is negative (suggesting that the person is a safer driver), then that needs to be discussed. Are you keeping the model even though it is counter intuitive? Why? The boss needs to know.


### 4. SELECT MODELS (25 Points)

Decide on the criteria for selecting the best multiple linear regression model and the best binary logistic regression model. Will you select models with slightly worse performance if it makes more sense or is more parsimonious? Discuss why you selected your models.


For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.? Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2 , (c) F-statistic, and (d) residual plots. For the binary logistic regression model, will you use a metric such as log likelihood, AIC, ROC curve, etc.? Using the training data set, evaluate the binary logistic regression model based on (a) accuracy, (b) classification error rate, (c) precision, (d) sensitivity, (e) specificity, (f) F1 score, (g) AUC, and (h) confusion matrix. Make predictions using the evaluation data set.

# Import Data

```{r, echo=FALSE}
url_git <- "https://raw.githubusercontent.com/GitableGabe/Data621_Data/main/"
```

```{r}
df_insur_eval <- 
  read.csv(paste0(url_git,"insurance-evaluation-data.csv"))

head(df_insur_eval)
```

```{r}
df_insur_train <- 
  read.csv(paste0(url_git,"insurance_training_data.csv"))

head(df_insur_train)
```

```{r}
dim(df_insur_train)
```
In the training dataset, there are 8,161 rows and 26 columns. We will remove the INDEX column because it is a unique identifier and will not be used.The two outcome variables are:
* TARGET_FLAG - a 0/1 variable that indicates if a insurance client has been in a car accident
* TARGET_AMT - a numeric variable that of insurance claim payout per car accident 

```{r}
df_insur_train <- df_insur_train %>% 
  select(-INDEX)
```

```{r}
df_insur_eval <- df_insur_eval %>% 
  select(-INDEX)
```


* There are 12 variables with discrete values and 13 variables with continuous values

##### DATA CLEANING


We noticed that there are characters in several of the columns that need to be cleaned up before the analysis. These will be removed and if necessary the variable will be converted to the appropriate data type.

```{r}
df_insur_train <- df_insur_train %>% 
  mutate(INCOME = gsub("\\$", "", INCOME), HOME_VAL = gsub("\\$", "", HOME_VAL),
         BLUEBOOK = gsub("\\$", "", BLUEBOOK), OLDCLAIM = gsub("\\$", "",
                                                               OLDCLAIM)) %>% 
  mutate(INCOME = gsub(",", "", INCOME), HOME_VAL = gsub(",", "", HOME_VAL),
         BLUEBOOK = gsub(",", "", BLUEBOOK), OLDCLAIM = gsub(",", "",
                                                             OLDCLAIM)) %>% 
  mutate(INCOME = as.numeric(INCOME), HOME_VAL = as.numeric(HOME_VAL),
         BLUEBOOK = as.numeric(BLUEBOOK), OLDCLAIM = as.numeric(OLDCLAIM))
```

```{r}
df_insur_train <- df_insur_train %>% 
  mutate(MSTATUS = gsub("z_","", MSTATUS), SEX = gsub("z_","", SEX),
         EDUCATION = gsub("z_","", EDUCATION), JOB = gsub("z_","", JOB),
         CAR_TYPE = gsub("z_","", CAR_TYPE), URBANICITY = gsub("z_","",
                                                               URBANICITY))
```

```{r}
df_insur_eval <- df_insur_eval %>% 
  mutate(INCOME = gsub("\\$", "", INCOME), HOME_VAL = gsub("\\$", "", HOME_VAL),
         BLUEBOOK = gsub("\\$", "", BLUEBOOK), OLDCLAIM = gsub("\\$", "",
                                                               OLDCLAIM)) %>% 
  mutate(INCOME = gsub(",", "", INCOME), HOME_VAL = gsub(",", "", HOME_VAL),
         BLUEBOOK = gsub(",", "", BLUEBOOK), OLDCLAIM = gsub(",", "",
                                                             OLDCLAIM)) %>% 
  mutate(INCOME = as.numeric(INCOME), HOME_VAL = as.numeric(HOME_VAL),
         BLUEBOOK = as.numeric(BLUEBOOK), OLDCLAIM = as.numeric(OLDCLAIM))
```

```{r}
df_insur_eval <- df_insur_eval %>% 
  mutate(MSTATUS = gsub("z_","", MSTATUS), SEX = gsub("z_","", SEX),
         EDUCATION = gsub("z_","", EDUCATION), JOB = gsub("z_","", JOB),
         CAR_TYPE = gsub("z_","", CAR_TYPE), URBANICITY = gsub("z_","",
                                                               URBANICITY))
```

* We will recode JOB into White Collar(Clerical, Doctor, Lawyer, Manager, and Professional), Blue Collar, and None (Student, Homemaker)
```{r}
df_insur_train <- df_insur_train %>% 
  mutate(JOB = ifelse(JOB=="Blue Collar", "Blue Collar", 
                      ifelse(JOB=="Student" | JOB=="Home Maker",
                             "None",
                             "White Collar")))
df_insur_eval <- df_insur_eval %>% 
  mutate(JOB = ifelse(JOB=="Blue Collar", "Blue Collar", 
                      ifelse(JOB=="Student" | JOB=="Home Maker",
                             "None", "White Collar")))
```

* We will also recode KIDSDRIV into a 0 or 1 (1+kids driving). Because there are a lot more insurance claims without kids dring than with kids driving.
```{r}
df_insur_train <- df_insur_train %>% 
  mutate(KIDSDRIV = ifelse(KIDSDRIV >= 1, 1, 0))

df_insur_eval <- df_insur_eval %>% 
  mutate(KIDSDRIV = ifelse(KIDSDRIV >= 1, 1, 0))
```

* Also, recode the yes/mo labels for marital status, parent status, red car, and revoked license variables as 1/0.

```{r}
df_insur_train <- df_insur_train %>% 
  mutate(MSTATUS = ifelse(MSTATUS == "No", "0", "1"),
         PARENT1 = ifelse(PARENT1 == "No", "0", "1"),
         RED_CAR = ifelse(RED_CAR == "no", "0", "1"),
         REVOKED = ifelse(REVOKED == "No", "0", "1"))
df_insur_eval <- df_insur_eval %>% 
  mutate(MSTATUS = ifelse(MSTATUS == "No", "0", "1"),
         PARENT1 = ifelse(PARENT1 == "No", "0", "1"),
         RED_CAR = ifelse(RED_CAR == "no", "0", "1"),
         REVOKED = ifelse(REVOKED == "No", "0", "1"))
```

* Lastly we will shorten the lables for Urbanicity and Turn Education into a factor with "< Highschool" as the reference variable.

```{r}
df_insur_train <- df_insur_train %>% 
  mutate(URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban",
                             "Urban", "Rural")) %>% 
  mutate(EDUCATION = factor(EDUCATION,levels = c("<High School",
                                                 "High School",
                                                 "Bachelors",
                                                 "Masters",
                                                 "PhD")))

df_insur_eval <- df_insur_eval %>% 
  mutate(URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban",
                             "Urban", "Rural")) %>% 
  mutate(EDUCATION = factor(EDUCATION,levels = c("<High School",
                                                 "High School",
                                                 "Bachelors",
                                                 "Masters",
                                                 "PhD")))
```

##### MISSING DATA ANS IMPUTATION


```{r}
#loop to count the NAs for each column
for (i in colnames(df_insur_train)){
  print(paste(i,"  ", sum(is.na(df_insur_train[,i])),sep = ""))
}
```
* There are NAs in three variable columns, 6 in AGE, 454 in YOJ (Years on the job) , and 510 in CAR_AGE. For these variable we will impute the median so as not to create an over fitting problem. Also, there was an irrational value of negative 3 for CAR_AGE, we replaced it with zero.

```{r}
df_insur_train <- df_insur_train %>% 
  mutate(AGE = ifelse(is.na(AGE),
                      median(AGE, na.rm = TRUE),
                      AGE), YOJ = ifelse(is.na(YOJ),
                                         median(YOJ, na.rm = TRUE), YOJ),
         CAR_AGE = ifelse(is.na(CAR_AGE),
                          median(CAR_AGE, na.rm = TRUE), CAR_AGE), 
         HOME_VAL = ifelse(is.na(HOME_VAL),
                           median(HOME_VAL,
                                  na.rm = TRUE), HOME_VAL),
         INCOME = ifelse(is.na(INCOME),
                         median(INCOME, na.rm = TRUE),
                         INCOME)) %>%
  mutate(CAR_AGE = ifelse(CAR_AGE < 0, 0, CAR_AGE))

summary(df_insur_train$CAR_AGE)
```
```{r}
#loop to count the NAs for each column
for (i in colnames(df_insur_eval)){
  print(paste(i,"  ", sum(is.na(df_insur_eval[,i])),sep = ""))
}
```
* There are NAs in five variable columns, 1 in AGE, 94 in YOJ (Years on the job) , 125 in INCOME, 111 HOME_VAL, and 129 in CAR_AGE. For these variable we will impute the median so as not to create an over fitting problem. 

```{r}
df_insur_eval <- df_insur_eval %>% 
  mutate(AGE = ifelse(is.na(AGE), median(AGE, na.rm = TRUE),
                      AGE), YOJ = ifelse(is.na(YOJ),
                                         median(YOJ, na.rm = TRUE), YOJ),
         CAR_AGE = ifelse(is.na(CAR_AGE), median(CAR_AGE, na.rm = TRUE),
                          CAR_AGE)) %>% 
  mutate(INCOME = ifelse(is.na(INCOME), median(INCOME,
                                               na.rm = TRUE), INCOME),
         HOME_VAL = ifelse(is.na(HOME_VAL),
                           median(HOME_VAL, na.rm = TRUE), YOJ))

summary(df_insur_eval$CAR_AGE)
```


###Exploratory Data analysis

Summary statistics for the numeric variables:
```{r}
df_insur_train %>% 
  select(TARGET_AMT, AGE, YOJ, INCOME, HOME_VAL, TRAVTIME, BLUEBOOK, TIF,
         OLDCLAIM, CLM_FREQ, MVR_PTS, CAR_AGE, HOMEKIDS) %>% 
  describe()
```
The skewness and Kurtosis values for the outcome variable TARGET_AMT strongly suggests that the distribution is likely not normal.

```{r}
df_insur_train %>% 
  select(TARGET_AMT, AGE, YOJ, INCOME, HOME_VAL, TRAVTIME, BLUEBOOK,
         TIF, OLDCLAIM, CLM_FREQ, MVR_PTS, CAR_AGE, HOMEKIDS) %>% 
  gather() %>% 
  ggplot(aes(x = value)) + 
  geom_histogram(fill = "cadetblue") + 
  facet_wrap(~key, scales = "free")
```

The histogram for TARGET_AMT, CAR_AGE, CLM_FREQ,HOME_VAL, INCOME, MRV_PTS, OLDCLAIM, and TIF are clearly not normally distributed and will need to be transformed if the residuals are not normally distributed.

We will explore the proportions of the discrete variables.

```{r}
df_insur_train %>% 
  select(TARGET_FLAG, KIDSDRIV, PARENT1, MSTATUS, SEX, EDUCATION,
         JOB, CAR_USE, CAR_TYPE, RED_CAR, REVOKED, URBANICITY) %>% 
  gather() %>% 
  ggplot(aes(x = value)) + 
  geom_bar(fill = "cadetblue") + 
  coord_flip()+
  facet_wrap(~key, scales = "free")
```

To check for coliniearity through the corellation of the variables

```{r}
mat <- df_insur_train %>% 
  select(-CAR_TYPE, -CAR_USE, -EDUCATION, -JOB, -SEX, -URBANICITY) %>% 
  mutate(PARENT1 = as.numeric(PARENT1), MSTATUS = as.numeric(MSTATUS),
         RED_CAR = as.numeric(RED_CAR), REVOKED = as.numeric(REVOKED)) %>% 
  cor()
corrplot(mat, method = "circle", diag = FALSE, order ="hclust", type = "lower")
```
* We do not seem to have very much concern for high collinearity at this point.

### Logistic Regression Model

```{r}
log_mod <- glm(TARGET_FLAG ~., data = df_insur_train[,-2],
               family = binomial(link = "logit"))
summary(log_mod)
```



```{r}
vif(log_mod)
```
* The degree of freedom adjusted variance inflation factors suggests that there is no concerning colinearity becasue all of the values are less than 3.

```{r}
log_step <- step(log_mod, direction = "backward", test = "LRT")
```

```{r}
summary(log_step)
```
The variable that positively impact the log odds of having car crash are the following:

* Kids driving
* Having kids at home (although this is a marginally significant p-value)
* Being a parent(vs not being a a parent)
* Having a longer travel time
* Having a car type other than minivan(when compared to minivan)
* Having an increased claims frequency 
* Having a revoked license
* Residing in an urban environment
* Having more points on the drivers license

The variable that negatively impact the log odds of having car crash are the following:

* Having a higher income
* Having a higher home value
* Being married
* Having a college of graduate level education as opposed to having less than a high school level education(there is no difference between having a high school diploma and not having one)
* Using the car for private as opposed to commercial use
* Having a higher Bluebook value for your vehicle
* Having a longer tenure as insurance client
* Having longer period of times between claims

### ASSESING MODEL PERFORMANCE

We are going to first predict the probabilities of a car crash using the final backward stepwise regression model from which we will then call the predicted car crash based on the probability of 0.5.

```{r}
df_insur_train$log_pred_prob <- predict(log_step,
                                        newdata = df_insur_train[,-c(1:2)],
                                        type = "response")
df_insur_train$log_pred <- ifelse(df_insur_train$log_pred_prob > 0.5, 1, 0)
```

Next we will assess model performance by calculating the area under the curve (AUC) for this model.

```{r}
pROC::auc(df_insur_train$TARGET_FLAG, df_insur_train$log_pred)
pROC::roc(df_insur_train$TARGET_FLAG~df_insur_train$log_pred,
          plot = TRUE, print.auc = TRUE)
```
The AUC of the model of .67 indicates that the model is only fair at predicting whether or not an insurance client will have a car crash.

We can get a clearer sense of how the model under performed by looking at a confusion matrix.

```{r}
confusionMatrix(as.factor(df_insur_train$log_pred),
                as.factor(df_insur_train$TARGET_FLAG), positive = "1")
```
* After fitting the final logistic model to the train data the accuracy obtained is 78.9%, but the sensitivity is extremely low at only 41% thus the balance accuracy is the same as the AUC at 66.8%. 
* It is worth noting that with such low sensitivity we can expect predictions to grossly under perform when predicting car crashes. 

#### PREDICTING CAR CRASHES WITH THE EVALUATIONS DATASET

```{r}
df_insur_eval$log_pred_prob <- predict(log_step,
                                       newdata = df_insur_eval[,-c(1:2)],
                                       type = "response")
df_insur_eval$log_pred <- ifelse(df_insur_eval$log_pred_prob > 0.5, 1, 0)
```

```{r}
df_insur_eval %>% 
  select(log_pred, KIDSDRIV, PARENT1, MSTATUS, EDUCATION, CAR_TYPE,
         REVOKED, URBANICITY, HOMEKIDS) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate_if(is.numeric, as.factor) %>% 
  pivot_longer(-log_pred, names_to = "key", values_to = "value") %>% 
  ggplot(aes(x = value, fill = log_pred)) + 
  geom_bar() + 
  scale_fill_discrete(labels = c("no crash", "crash"),
                      name = "Predicted Outcome") +
  coord_flip()+
  facet_wrap(~key, scales = "free")
```


```{r}
df_insur_eval %>% 
  select(log_pred, INCOME, HOME_VAL, TRAVTIME, TIF, OLDCLAIM,
         CLM_FREQ, MVR_PTS, BLUEBOOK) %>% 
  pivot_longer(-log_pred, names_to = "key", values_to = "value") %>%  
  ggplot(aes(y = value, x = as.factor(log_pred), fill = as.factor(log_pred))) + 
  geom_boxplot() + 
  scale_fill_discrete(labels = c("no crash", "crash"),
                      name = "Predicted Outcome") +
  xlab("Predicted Outcome")+
  facet_wrap(~key, scales = "free")
```

Assessing the predicted car crashes for the evaluation dataset, seems to largely reflect what put into the model. Areas with stronger predictions were:

* Being a parent(vs not being a a parent)
* Having a longer travel time
* Having a car type other than minivan
* Having an increased claims frequency 
* Having a revoked license
* Residing in an urban environment
* Having a lower Bluebook value for your vehicle

We do not see any change in the predicted car crashes with respect to the variable home values. 

#### MULTIPLE LINEAR REGRESSION

```{r}
mlr_mod <- lm(TARGET_AMT ~., data = df_insur_train[,-c(1,26:27)])
summary(mlr_mod)
```

```{r}
vif(mlr_mod)
```
* The degree of freedom adjusted variance inflation factors suggests that there is no concerning colinearity because all of the values are less than 3.

```{r}
mlr_step <- step(mlr_mod, direction = "backward", test = "F")
```

* Note that although bluebook was not dropped from the model it is not significant, thus we will drop from the final model. After dropping bluebook from the model, the variables of sex, old claim, and kids at home were no longer significant, thus these too were also dropped from the model.

```{r}
mlr_final <- lm(TARGET_AMT ~ KIDSDRIV + INCOME + PARENT1 + MSTATUS +
                  TRAVTIME + CAR_USE + TIF + CAR_TYPE + CLM_FREQ + REVOKED +
                  MVR_PTS + CAR_AGE + URBANICITY,
                data = df_insur_train[,-c(1, 26:27)])
```


```{r}
summary(mlr_final)
```

The variable that positively impact the average cost of having car crash are the following:

* Kids driving
* Being a parent(vs not being a a parent)
* Having a longer travel time
* Having a car type other than minivan(when compared to minivan)
* Having an increased claims frequency 
* Having a revoked license
* Residing in an urban environment
* Having higher points on drivers license 
* 

The variable that negatively impact the average cost of having crash are the following:

* Having a higher income
* Being married
* Using the car for private as opposed to commercial use
* Having a longer tenure as insurance client
* Having an older car

### TEST MODEL ASSUMPTIONS

```{r}
plot(mlr_final)
```
1. Linearity - the first plot shows that the relationship between target amount and the predictor variables in the final model is linear, so the assumption is net.
2. Normality - the second plot shows that there is an approximate normal distribution of the residuals in the final model.
3. Equality of Variances - the third plot that there are some unequal variance, however, the relationship is largely homoscedastic.
4. Leverage / High Influence - the fourth plot shows that there are a few outliers with very high claim, but they do not violate cooks distance.

* Thus we can trust the results of the final model.

### ASSESING MODEL PERFORMANCE

We are going to first predict the amount of the crash using the final model, we will then calculate the RMSE using the predictions.

```{r}
df_insur_train$mlr_pred <- predict(mlr_final,
                                   newdata = df_insur_train[,-c(1:2, 26:27)],
                                   type = "response")
RMSE(df_insur_train$mlr_pred, df_insur_train$TARGET_AMT)
summary(mlr_final)$adj.r.squared
```
* The RMSE for this model suggest an average deviation in the predicted claim amount from the true claim amount of $4,545. This suggests that the model is not doing a particularly good job at predicting accurate claim amounts. This is not surprising given that the R squared of the final model only explain 6.4% of the total variation in the claim amount. 

#### PREDICTING AMOUNT OF CLAIM FOR CAR CRASHES WITH THE EVALUATIONS DATASET

```{r}
df_insur_eval$mlr_pred <- predict(mlr_final,
                                  newdata = df_insur_eval[,-c(1:2, 26:27)],
                                  type = "response")
```

```{r}
coef(mlr_final)
```

```{r}
df_insur_eval %>% 
  select(mlr_pred, INCOME, HOME_VAL, TRAVTIME, TIF, CLM_FREQ, MVR_PTS,
         CAR_AGE) %>%
  mutate_if(is.character, as.numeric) %>% 
  pivot_longer(-mlr_pred, names_to = "key", values_to = "value") %>%  
  ggplot(aes(x = value, y = mlr_pred)) + 
  geom_point() + 
  geom_smooth() +
  ylab("Predicted Claim Amount USD")+
  facet_wrap(~key, scales = "free")
```

```{r}
df_insur_eval %>% 
  select(mlr_pred, KIDSDRIV, MSTATUS, CAR_TYPE, REVOKED, URBANICITY, CAR_USE,
         PARENT1) %>% 
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(-mlr_pred, names_to = "key", values_to = "value") %>% 
  mutate(mlr_pred = as.numeric(mlr_pred)) %>% 
  ggplot(aes(x = mlr_pred, y = value)) + 
  geom_boxplot(fill = "cadetblue") + 
  xlab("Predicted Claim Amount USD") +
  facet_wrap(~key, scales = "free")
```